---
title: "CCA"
output: html_document
date: "2024-04-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(CCA)
library(CCP)
library(ggplot2)
library(ggpubr)
```

## Load data

```{r}

dataframe <- read.csv("/data/HR_data.csv")
dataframe <- na.omit(dataframe)
col_ <- colnames(dataframe)

xcols <- c(col_[2:56], col_[58])
ycols_all <- c(col_[57], col_[59:68])
ycols_p <- c(col_[61], col_[63], col_[65], col_[67], col_[68])            #Negative emotions
ycols_n <- c(col_[57], col_[59], col_[60], col_[62], col_[64], col_[66])  #Positive emotions

x <- dataframe[xcols]
y_p <- dataframe[ycols_p]
y_n <- dataframe[ycols_n]

# one hot encoding
x$Cohort <- as.numeric(as.factor(x$Cohort)) 
x$Phase <- as.numeric(as.ordered(x$Phase))
x$Round <- as.numeric(as.ordered(x$Round))
```

# CCA positive emotions

## CCA and hypothesis test (Positive)

```{r}
cca_p <- cc(x, y_p)
cat("Correlation for positive emotions:\n ", cca_p$cor)

cat("\n\nDimensions for U for positive emotions: ", dim(cca_p$ycoef) )
cat("\nDimensions for V for positive emotions: ",dim(cca_p$xcoef))

#Hypothesis test
rho_p <- cca_p$cor
N <- 307
p <- length(x)
q_p <- length(y_p)

cat("\n\nHypothesis test for positive emotions: \n")
p.asym(rho_p, N, p, q_p)
```

Only the two first components are significant

## CCA for the first 2 components (Positive)

You can also embed plots, for example:

```{r}
v1_p <- cca_p$xcoef[0:56, 1]
v2_p <- cca_p$xcoef[0:56, 2]

u1_p <- cca_p$ycoef[0:q_p, 1]
u2_p <- cca_p$ycoef[0:q_p, 2]
```

## Plot for V for positive emotions (Positive)

```{r}
df_v1_p <- data.frame(x_feature = c(1:56), coefficient = v1_p, component = "v1_p")
df_v2_p <- data.frame(x_feature = c(1:56), coefficient = v2_p, component = "v2_p")

# plotting the largest N components
n1_p <- 5
index1_p <- sort(abs(v1_p), index.return = TRUE, decreasing = TRUE)$ix[n1_p + 1:56]
col1_p <- colnames(x)
col1_p[index1_p] <- ""

# plotting the first component
v1_plot_p <- ggplot(data = df_v1_p, aes(x = x_feature, y = coefficient)) +
  geom_point(size = 4) +
  geom_segment(aes(x = x_feature, xend = x_feature, y = 0, yend = coefficient),
               linewidth = 2, alpha = 0.5) +
  theme(text = element_text(size = 20),
        legend.text = element_text(size = 20)) +
  geom_text(aes(x = x_feature, y = coefficient, label = col1_p),
            fontface = "bold", color = "red")

# plotting the largest N components
n2_p <- 6
index2_p <- sort(abs(v2_p), index.return = TRUE, decreasing = TRUE)$ix[n2_p + 1:56]
col2_p <- colnames(x)
col2_p[index2_p] <- ""

# plotting the second component
v2_plot_p <- ggplot(data = df_v2_p, aes(x = x_feature, y = coefficient)) +
  geom_point(size = 4) +
  geom_segment(aes(x = x_feature, xend = x_feature, y = 0, yend = coefficient),
               linewidth = 2, alpha = 0.5) +
  theme(text = element_text(size = 20),
        legend.text = element_text(size = 20)) +
  geom_text(aes(x = x_feature, y = coefficient, label = col2_p),
            fontface = "bold", color = "red")


ggarrange(v1_plot_p, v2_plot_p, labels = c("v1_p", "v2_p"),
          ncol = 1, nrow = 2,
          font.label = list(size = 20, color = "black", face = "bold"))



```

## Plot for U for positive emotions (Positive)

```{r}
df_u1_p <- data.frame(y_feature = c(1:q_p), coefficient = u1_p, component = "u1_p")
df_u2_p <- data.frame(y_feature = c(1:q_p), coefficient = u2_p, component = "u2_p")

col_p <- colnames(y_p)

u1_plot_p <- ggplot(data = df_u1_p, aes(x = y_feature, y = coefficient)) +
  geom_point(size = 4) +
  geom_segment(aes(x = y_feature, xend = y_feature, y = 0, yend = coefficient),
               linewidth = 2, alpha = 0.5) +
  theme(text = element_text(size = 20),
        legend.text = element_text(size = 20)) +
  geom_text(aes(x = y_feature, y = coefficient, label = col_p),
            fontface = "bold", color = "red", size = 7)

u2_plot_p <- ggplot(data = df_u2_p, aes(x = y_feature, y = coefficient)) +
  geom_point(size = 4) +
  geom_segment(aes(x = y_feature, xend = y_feature, y = 0, yend = coefficient),
               linewidth = 2, alpha = 0.5) +
  theme(text = element_text(size = 20),
        legend.text = element_text(size = 20)) +
  geom_text(aes(x = y_feature, y = coefficient, label = col_p),
            fontface = "bold", color = "red", size = 7)


ggarrange(u1_plot_p, u2_plot_p, labels = c("u1_p", "u2_p"),
          ncol = 1, nrow = 2,
          font.label = list(size = 20, color = "black", face = "bold"))

```

# CCA for negative emotions

## CCA and hypothesis test (Negative)

```{r}
cca_n <- cc(x, y_n)
cat("Correlation for positive emotions:\n ", cca_n$cor)

cat("\n\nDimensions for U for positive emotions: ", dim(cca_n$ycoef) )
cat("\nDimensions for V for positive emotions: ",dim(cca_n$xcoef))

#Hypothesis test
rho_n <- cca_n$cor
N <- 307
p <- length(x)
q_n <- length(y_n)

cat("\n\nHypothesis test for negative emotions: \n")
p.asym(rho_n, N, p, q_n)
```

Only the two first components are significant

## CCA for the first 2 components (Negative)

```{r}
v1_n <- cca_n$xcoef[0:56, 1]
v2_n <- cca_n$xcoef[0:56, 2]

u1_n <- cca_n$ycoef[0:q_n, 1]
u2_n <- cca_n$ycoef[0:q_n, 2]
```

## Plot for V for positive emotions (Negative)

```{r}

# ADVANCED PLOTTING, REQUIRES SOME PACKAGES
df_v1_n <- data.frame(x_feature = c(1:56), coefficient = v1_n, component = "v1_n")
df_v2_n <- data.frame(x_feature = c(1:56), coefficient = v2_n, component = "v2_n")

# plotting the largest N components
n1_n <- 5
index1_n <- sort(abs(v1_n), index.return = TRUE, decreasing = TRUE)$ix[n1_n + 1:56]
col1_n <- colnames(x)
col1_n[index1_n] <- ""

# plotting the first component
v1_plot_n <- ggplot(data = df_v1_n, aes(x = x_feature, y = coefficient)) +
  geom_point(size = 4) +
  geom_segment(aes(x = x_feature, xend = x_feature, y = 0, yend = coefficient),
               linewidth = 2, alpha = 0.5) +
  theme(text = element_text(size = 20),
        legend.text = element_text(size = 20)) +
  geom_text(aes(x = x_feature, y = coefficient, label = col1_n),
            fontface = "bold", color = "red")

# plotting the largest N components
n2_n <- 6
index2_n <- sort(abs(v2_n), index.return = TRUE, decreasing = TRUE)$ix[n2_n + 1:56]
col2_n <- colnames(x)
col2_n[index2_n] <- ""

# plotting the second component
v2_plot_n <- ggplot(data = df_v2_n, aes(x = x_feature, y = coefficient)) +
  geom_point(size = 4) +
  geom_segment(aes(x = x_feature, xend = x_feature, y = 0, yend = coefficient),
               linewidth = 2, alpha = 0.5) +
  theme(text = element_text(size = 20),
        legend.text = element_text(size = 20)) +
  geom_text(aes(x = x_feature, y = coefficient, label = col2_n),
            fontface = "bold", color = "red")



ggarrange(v1_plot_n, v2_plot_n, labels = c("v1_n", "v2_n"),
          ncol = 1, nrow = 2,
          font.label = list(size = 20, color = "black", face = "bold"))
```

## Plot for U for positive emotions (Negative)

```{r}

df_u1_n <- data.frame(y_feature = c(1:q_n), coefficient = u1_n, component = "u1_n")
df_u2_n <- data.frame(y_feature = c(1:q_n), coefficient = u2_n, component = "u2_n")

col_n <- colnames(y_n)


u1_plot_n <- ggplot(data = df_u1_n, aes(x = y_feature, y = coefficient)) +
  geom_point(size = 4) +
  geom_segment(aes(x = y_feature, xend = y_feature, y = 0, yend = coefficient),
               linewidth = 2, alpha = 0.5) +
  theme(text = element_text(size = 20),
        legend.text = element_text(size = 20)) +
  geom_text(aes(x = y_feature, y = coefficient, label = col_n),
            fontface = "bold", color = "red", size = 7)



u2_plot_n <- ggplot(data = df_u2_n, aes(x = y_feature, y = coefficient)) +
  geom_point(size = 4) +
  geom_segment(aes(x = y_feature, xend = y_feature, y = 0, yend = coefficient),
               linewidth = 2, alpha = 0.5) +
  theme(text = element_text(size = 20),
        legend.text = element_text(size = 20)) +
  geom_text(aes(x = y_feature, y = coefficient, label = col_n),
            fontface = "bold", color = "red", size = 7)


ggarrange(u1_plot_n, u2_plot_n, labels = c("u1_n", "u2_n"),
          ncol = 1, nrow = 2,
          font.label = list(size = 20, color = "black", face = "bold"))
```
